A Type System for Safe and Scalable Parallelism using Logical Regions \\ (Extended Version)

### 2012-07-19 doc/popl2013full/langdef.tex
connected dots in E-Pack proof

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these
properties.\footnote{Note that in the definition of
$\mapconsist{\Omega}$, the quantified variables $r_i$ and $r_j$
range over both logical and physical regions.  Since $M$'s domain is
logical regions, we extend $M$ to constraints with mixed logical and
physical regions in the standard way by defining
$M \llbracket \rho \rrbracket = \rho$.}

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\hat \Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\hat \Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

@@

The only thing remaining to show is $\valueconsist{v'}{T_1}$.
From [E-Pack] we have   $v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle$.
From [T-Pack] we have $T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1$
and $\Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^*$.

Observe that 
$$M \llbracket \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \rrbracket$$
replaces all free region names in $T_2$ and particularly $\Omega_1$ by physical regions, so
that the only region names remaining are $r'_1,\ldots,r'_n$.  It is then easy to show that
$$\mapconsist[\{(r'_i,\rho_i)\}]{M \backslash \{r'_1,\ldots, r'_n \}\llbracket \Omega_1 \rrbracket}$$
from which $\valueconsist{v'}{T_1}$ follows (see Figure~\ref{fig:constprop}).

>>>>

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these
properties.\footnote{Note that in the definition of
$\mapconsist{\Omega}$, the quantified variables $r_i$ and $r_j$
range over both logical and physical regions.  Since $M$'s domain is
logical regions, we extend $M$ to constraints with mixed logical and
physical regions in the standard way by defining
$M \llbracket \rho \rrbracket = \rho$.  And although the definition of
$\traceconsist{E}$ refers to $M$ and $\Gamma$, the restriction that reduction
functions must be monomorphic and visible in all functions means that 
$\traceconsist{E}$ will be insensistive to the particular $M$ and $\Gamma$.}

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\hat \Omega} & \Leftrightarrow & (\valueconsist[]{v}{\hat T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\hat \Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

\begin{lem}
\label{lemma:mappedomega}
\rm
$\mapconsist{\Omega}$ if and only if $\mapconsist[\emptyset]{M \llbracket \Omega \rrbracket}$.
\end{lem}

@@

The only thing remaining to show is $\valueconsist{v'}{T_1}$.
From [E-Pack] we have   $v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle$ and $M(r_i) = \rho_i$.
From [T-Pack] we have $T_1 = \exists r'_1, \ldots, r'_n.\ T_2\text{ where }\Omega_1$
and $\Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^*$.

Let $M \llbracket T_1 \rrbracket = \exists r'_1, \ldots, r'_n.\ \hat T_2\text{ where }\hat \Omega_1$.
We require that $\valueconsist[]{v}{\hat T_2[\rho_1/r'_1,\ldots,\rho_k/r'_k]}$ and $\mapconsist[\{(r'_i,\rho_i)\}]{\hat \Omega_1}$.
The right hand side of the first condition can be rewritten as $\hat T_2[M(r_1)/r'_1,\ldots,M(r_k)/r'_k]$, which is equal
to $M \llbracket T_2 [r_1/r'_1,\ldots,r_k/r'_k] \rrbracket$.

Similarly, observe that
$$\hat \Omega_1 [ \rho_1/r'_1,\ldots,\rho_k/r'_k ] = M \llbracket \Omega_1 [ r_1/r'_1,\ldots,r_k/r'_k ] \rrbracket$$
.  With two applications of Lemma~\ref{lemma:mappedomega}, we have $\mapconsist[\{(r'_i,\rho_i)\}]{\hat \Omega_1}$
if and only if $\mapconsist{\Omega_1 [ r_1/r'_1,\ldots,r_k/r'_k ]}$.  The latter constraints are a subset of
$\Omega^*$, and must be satisfied if $\mapconsist{\Omega^*}$, which we have from Lemma~\ref{lemma:omegaclosure}.

\begin{comment}
For the second condition, consider each constraint $x_i \leq x_j \in \hat \Omega_1$, where each $x_i$ is either a 
physical region $\rho_i$ or one of the bounded logical regions $r'_i$.

We must therefore show $\valueconsist{v}{T_2[\rho_1/r'_1,\ldots,\rho_k/r'_k]}$ and $\mapconsist[\{(r_i',\rho_i)\}]{M \llbracket \Omega_1 \rrbracket}$.  We have $\rho_i = M(r_i)$ from [E-Pack] and a partial application of $M$ satisfies
the first condition.  To show the second condition, we apply the renaming $[r_1/r'_1,\ldots,r_k/r'_k]$ to both sides,
yielding
$$\mapconsist[\{(r_i,\rho_i)\}]{M \llbracket \Omega_1 [r_1/r'_1,\ldots,r_k/r'_k] \rrbracket}$$
,  Observe that the left hand side is a subset of $M$
By partially applying $M$ to $T_2[r_1/r'_1,\ldots,r_k/r'_k]$, we get $\valueconsist{v}{T_2[\rho_1/r'_1,\ldots,\rho_k/r'_k]}$, satisfying the first half o
Observe that 
$$M \llbracket \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \rrbracket$$
replaces all free region names in $T_2$ and particularly $\Omega_1$ by physical regions, so
that the only region names remaining are $r'_1,\ldots,r'_n$.  It is then easy to show that
$$\mapconsist[\{(r'_i,\rho_i)\}]{M \backslash \{r'_1,\ldots, r'_n \}\llbracket \Omega_1 \rrbracket}$$
from which $\valueconsist{v'}{T_1}$ follows (see Figure~\ref{fig:constprop}).
\end{comment}
